---
title: "Prediction of exercising"
author: "Rafael Lopez"
date: "11/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction 
We are analizing the Gym data of how well the people exercise.We want to create a predicted model to correctly classify how well people perform in the gym.We use Random Forest model to create a very accurate prediction.Half of the data was used to create the model and the other half was used to validate the data.

##Loading 
We loaded in the two part of the data, training and testing. We summarize the data and clean all the unnecessary information.
```{r,echo=TRUE}
#download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv","lift_train.csv",method = "curl")
lift_train <- read.csv("lift_train.csv",na.strings = c("NA","","#DIV/0!"),stringsAsFactors = FALSE)
#download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv","lift_test.csv",method = "curl")
lift_test<-read.csv("lift_test.csv",na.strings = c("NA","","#DIV/0!"),stringsAsFactors = FALSE)
##### Cleaning data ####
library(caret)
library(rpart)

### get rid of zero and incomplete data (cleaning) ####
colSums(lift_train == 0 | is.na(lift_train)) ### if is zero or na = true
#number of real observation per column
liftreal<-colSums(!(lift_train == 0 | is.na(lift_train))) ### if is zero or na = false
which(liftreal== 0) ## finding not real data.
completecols  <-which(liftreal > 12000) #### more complete data per columns
### clean up data frame ####
lift_train2 <- lift_train[,completecols]
lift_train2[,1:7]<-list (NULL) ## get rid of columns 1:7 not useful data
lift_train2$classe <- as.factor(lift_train2$classe)#changing from chr to Factor

```
## Centering the data
We center and scale the data to form a normal distribution.
```{r,echo=TRUE}

lift_train3 <-scale(lift_train2[-53],center = TRUE, scale = TRUE)
lift_train3 <- as.data.frame(lift_train3)
lift_train3$classe <- lift_train2$classe

```
# Data Partission 
We create a data partission to have information for training and validation.
```{r,echo=TRUE}
lift_trainpartisian <- createDataPartition(lift_train3$classe, list = FALSE)
lift_train4 <- lift_train3[lift_trainpartisian,]
lift_test4 <- lift_train3[-lift_trainpartisian,]
```
# Building the Decision Tree
We use the Random Forest for building the predictive model.

```{r,echo=TRUE}
library("caret")
library("randomForest")
library("rpart")

set.seed(123456)

fitliftrf8 <- train(classe ~.,method = "rf",data = lift_train4)
```
# Cross Validation
We performed the cross validation with the validation data from the partissian.Also we use confusion Matrix to confirm accuracy.
```{r,echo=TRUE}
predict8 <-predict(fitliftrf8,newdata = lift_test4)
confurf8 <- confusionMatrix(predict8,lift_test4$classe)
confurf8
```

# Predicting test data
The prediction model allow us to predict the 20 test cases.
```{r, echo = TRUE}
predict4 <- predict(fitliftrf8,lift_test)
predict4 
```




